### 2023-12-11 manuscript/paper.Rmd
update

---
title: "Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population"
thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
#subtitle: "Short title goes here!"
author:
- Amos Okutse, Brown University
- Monic Colon-Vargas, Brown University
abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.

**Methods:** Proin tellus orci, posuere sed cursus at, bibendum ac odio. Nam consequat non ante eget aliquam. Nulla facilisis tincidunt elit. Nunc hendrerit pellentesque quam, eu imperdiet ipsum porttitor ut. Interdum et malesuada fames ac ante ipsum primis in faucibus. Suspendisse potenti. Duis vitae nibh mauris. Duis nec sem sit amet ante dictum mattis. Suspendisse diam velit, maximus eget commodo at, faucibus et nisi. Ut a pellentesque eros, sit amet suscipit eros. Nunc tincidunt quis risus suscipit vestibulum. Quisque eu fringilla massa."
keywords: "pandoc, r markdown, knitr"
date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
#mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
bibliography: bibfile.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

```{r source-file, echo=FALSE, include=FALSE}
# source file 
source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")

#source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

# Introduction

The clinical importance of systolic blood pressure (SBP) stands as a
cornerstone extensively acknowledged and documented across the continuum
of medical literature's data collection [@gurven2012does;
@su2022association]. Elevated SBP stands as a precursor linked with a
spectrum of critical health conditions, encompassing cardiovascular
diseases, stroke, and renal impairment, among others. This nexus between
heightened SBP and these significant health adversities underscores the
gravity of SBP as a predictive marker for adverse health outcomes.
Noteworthy is the pivotal concern surrounding the trajectory of SBP in
the natural aging process, a trajectory observed ubiquitously across
diverse populations.

In western societies aged 40 years and above, a demonstrable pattern
emerges, revealing an approximate elevation of 7 mmHg in SBP per decade
among individuals above 40 years. This discernible and consistent
increment in SBP with advancing age accentuates its profound impact
within the broader context of aging-related health dynamics. Similarly,
while diastolic blood pressure (DBP), demonstrates a concurrent rise
correlating with age, presents a substantially lower rate of increase
when juxtaposed against the ascending trajectory observed in SBP
[@gurven2012does; @su2022association]. Thus, in light of the
significance attached to SBP dynamics within aging populations and the
intricate interplay between age, sleep, cholestrol levels and blood
pressure alterations, our investigative focus is oriented towards the
exploration of variables that exert potential influence on systolic
blood pressure (SBP). This deliberate focus aims to elucidate the
multifaceted nature of factors contributing to SBP variability, enabling
a more comprehensive understanding of its determinants within the
framework of health and aging.

```{=html}
<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized
contributor to systolic blood pressure (SBP), with individuals
exhibiting higher BMI values demonstrating a propensity for elevated
SBP. Gender also holds a discernible association with SBP dynamics.
Empirical evidence in CITE demonstrates that women typically manifest
lower SBP than men until the onset of menopause, after which women
exhibit higher SBP levels compared to men. Additionally, heightened
levels of cholesterol are closely linked to vascular damage. The
accumulation of cholesterol deposits and calcium within arteries
precipitates arterial stiffening, necessitating augmented cardiac
efforts to circulate blood, potentially culminating in elevated blood
pressure. This study endeavors to examine the interrelationships between
age and cholesterol levels, emphasizing consideration of potential
confounding variables. Table 1 comprehensively delineates the array of
variables under scrutiny, ensuring a comprehensive analysis of their
collective impact. -->
```
# Methods

## Study population

Data utilized in this paper is sourced from the National Health and
Nutrition Examination Survey (NHANES), a comprehensive nationwide survey
administered by the National Center for Health Statistics (NCHS) via the
Centers for Disease Control and Prevention (CDC). The survey assesses
the health and nutrition of the entire non-institutionalized US
population, spanning all ages and residing in all 50 states as well as
Washington D.C. As such, the survey provides a cross-sectional view of a
representative sample of the US population. Further information about
NHANES can be found at www.cdc.gov/nchs/nhanes.

## Data

Our current analyses combine the 2015 - 2018 NHANES survey cycles to
yield $n = 19225$ observations on 35 covariates. This sample size was
comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and
2017/2018 survey cycles, respectively. Analyses excluded individuals
with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055),
and body mass index (BMI) or those with BMI \> 25 kg/$m^2$ (overweight)
(n = 5521). Individuals that reported being on anti-hypertensive
medication were also excluded from further analyses (n = 2944). Our
final analyses were based on a sample of n = 1977 observations on 28
covariates.

### Outcome definition:

We defined our outcome as systolic and diastolic blood pressure. These
variables are measured by trained examiners using standardized
procedures. Given that systolic and diastolic blood pressure
measurements are taken at least four times on an individual, our
definition of these outcome is based on an average of the first three
blood pressure measurements.

```{=html}
<!-- 
### Exposures:

Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on weekdays or
workdays". We then categorized this variable into three groups, that is,
\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
group in our analysis. We also explored the association between total
cholestrol (in mmol/L) level on systolic blood pressure including
potential effect modification by other factors.
-->
```
### Covariates:

```{=html}
<!-- Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on weekdays or
workdays". We then categorized this variable into three groups, that is,
\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
group in our analysis. -->
```
Our analyses included the following as covariates: Race divided into
four groups as Mexican American, white, black and other race. Alcohol
consumption was grouped into drinking, no drinking, not recorded.
Smoking status as smoking, not smoking, not recorded. Diabetes was
defined as yes, no, borderline, or not recorded. Hypertension was
defined as yes, no, or not recorded. Snorting was defined as yes, no,
and not recorded. US citizenship status was defined as citizen by birth
or naturalization, don't know, not a citizen, or refused to answer.
Education level was grouped into four categories including graduate
studies, high school, less than grade 12 or some college. Additional
covariates included marital status, gender, age, albumin,
creatinine,hemoglobin, total cholesterol (TC), aspertate
aminotransferase (AST), high-density lipoprotein (HDL), and body mass
index (BMI). Details about these variables can be found at
<https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx>. Age, albumin,
creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as
continuous variables whereas gender, alcohol consumption, diabetes,
smoking, race, hypertension, and snorting were analyzed as categorical
variables. Sleep duration on workdays was also recorded and was
categorized into three groups, that is, \< 6 h, 6--8 h, ≥ 8 h
respectively, and used \< 6 h as the reference group in our analysis.
Table \@ref{tab:tabone} highlights the variable names and descriptions
as utilized in this study.

```{=tex}
\begin{table}[H]

\caption{\label{tab:tabone}Variable descriptions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Name} & \textbf{Description}\\
\midrule
\textbf{SEQN} & \textbf{sequence number} & Respondent number\\
\textbf{SDMVPSU} & \textbf{psu} & Masked variance unit pseudo-PSU variable for variance estimation\\
\textbf{WTINT2YR} & \textbf{weights} & Full sample 2-year interview weights\\
\textbf{SDMVSTRA} & \textbf{strata} & Masked variance unit pseudo-stratum variable for variance estimation\\
\textbf{RIAGENDR} & \textbf{gender} & Respondent's number\\
\addlinespace
\textbf{RIDAGEYR} & \textbf{age (yrs)} & Respondent's age in years\\
\textbf{DMDMARTL} & \textbf{marital status} & Marital status\\
\textbf{INDFMIN2} & \textbf{income category} & Total family income (reported as a range value in dollars)\\
\textbf{RIDRETH3} & \textbf{race} & Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category\\
\textbf{DMDHHSZA} & \textbf{children <5} & Number of children aged 5 years or younger in the household\\
\addlinespace
\textbf{DMDEDUC2} & \textbf{education level} & What is the highest grade or level of school\{you have/SP has\} completed or the highest degree \{you have/s/he has\} received?\\
\textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on weekdays or workdays\\
\textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
\textbf{ALQ121} & \textbf{alcohol use} & In the past 12 months, how often did you drink any type of alcoholic beverage?\\
\addlinespace
\textbf{LBDHDDSI} & \textbf{hdl} & High density lipoprotein\\
\textbf{LBDSALSI} & \textbf{albumin} & Albumin (g/L)\\
\textbf{DIQ010} & \textbf{diabetes} & Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?\\
\textbf{BPQ020} & \textbf{hypertension} & \{Have you/Has SP\} ever been told by a doctor or other health professional that \{you/s/he\} had hypertension, also called high blood pressure?\\
\textbf{URXCRS} & \textbf{creatinine} & Creatinine, urine (umol/L)\\
\addlinespace
\textbf{SLQ040} & \textbf{snort} & In the past 12 months, how often did \{you/SP\} snort, gasp, or stop breathing while \{you were/s/he was\} asleep?\\
\textbf{LBDTCSI} & \textbf{total cholestrol levels} & Total cholesterol (mmol/L)\\
\textbf{LBXHGB} & \textbf{hemoglobin} & Hemoglobin (g/dL)\\
\textbf{LBXSASSI} & \textbf{AST} & Aspertate aminotransferase\\
\textbf{SMQ040} & \textbf{smoke} & Do you now smoke cigarettes?\\
\addlinespace
\textbf{BPXDI} & \textbf{blood pressure} & Systolic and diastolic blood pressure taken as the average of the first three measurements\\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabone, echo=FALSE, eval=FALSE}
tab.one <- vars %>% 
  kable(format = "latex", caption = "Variable descriptions", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("HOLD_position")) %>%
  row_spec(c(0), bold = TRUE) %>%
  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>%
  column_spec(3, width = "8cm")
tab.one
```

## Statistical modeling

With the purpose of identifying influential variables impacting blood
pressure, a variable selection process was conducted. The initial
approach involved employing best subset selection with 10-fold
cross-validation to ascertain the optimal number of subsets. Notably,
for systolic blood pressure, the forward best subset selection method
discerned 13 variables, whereas diastolic blood pressure exhibited 19
selected variables. However, a limitation inherent in forward best
subset selection is its inability to eliminate previously selected
features, potentially disregarding their relevance in light of newly
added variables. Consequently, this methodology may yield sub-optimal
variable selections due to its lack of adaptability.

In an effort to mitigate this limitation and introduce regularization,
LASSO regularization was employed. This involved a 10-fold
cross-validation process to determine the optimal penalty parameter,
$\lambda$ . For diastolic blood pressure, the $\lambda$ value associated
with the lowest mean squared error (MSE) led to the selection of 25
variables. Conversely, when considering systolic blood pressure, the
$\lambda$ value minimizing the MSE resulted in the exclusion of a sole
variable. Subsequently, a $\lambda$ value was strategically chosen to
ensure a negligible increase of no more than 1% in MSE, ultimately
leading to the inclusion of 29 variables in the model.

# Results

## Descriptive statistics

Table \@ref(tab:tabtwo) summarizes participant characteristics in the
2015 - 2018 NHANES survey cycles stratified by gender. The 2015-2016
survey cycle comprised 53.77% of the total analyzed sample whereas the
2017 - 2018 survey cycle comprised 46.23% of the total analyzed sample.
Among the participants, the proportion of males and females were 53.76%
(n = 1063) and 46.23% (n = 914), respectively. With race, the proportion
of Mexican American, White and Black were 9.16%, 34.90% and 17.10%,
respectively. Overall, the mean (SD) values for age, albumin, SBP, DBP,
hemoglobin, TC, AST, HDL, BMI were 43.79 (17.43) years, 43.04 (3.59)
g/L, 118.61 (16.41) mmHg, 69.60 (11.18) mmHg, 13.94 (1.46)g/dL, 4.81
(1.02) mmol/L, 23.47 (13.17) IU/L, 1.60 (0.43) mmol/L, and 22.08 (2.03)
kg/m2, respectively. Among the participants, 73.70% were alcohol
drinkers, 5.06% were diabetic, 7.79% were hypertensive, 13.76%
experienced snorting whereas 21.60% were smokers. Sleep duration was
divided into three groups, which were \< 6 h, 6--8 h, ≥ 8 h, each with a
proportion of 7.89%, 40.06%, 52.05%, respectively. Male and female
participants differed significantly on all characteristics except
citizenship status, and the cycle of the survey.

```{r tabtwo, echo=FALSE, message=FALSE, warning=FALSE}
tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase. SBP = Systolic blood pressure, DBP = Diastolic blood pressure.", 
                 longtable = TRUE)
```

## Univariate regression analysis

Table \@ref(tab:tabthree) presents the summary of univariate regression
analysis results exploring the association between the exposures,
selected potential confounding variables and systolic blood pressure. We
found substantial associations between ...

```{r tabthree, echo=FALSE}
## univariate analysis
sbp.univariate %>% 
  as_kable_extra(format = "latex", booktabs = TRUE, digits = 2, caption = "Univariate analyses of exposures, potential confounding factors and systolic blood pressure. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase.", longtable = TRUE)
```

## Multivariable Regression Analysis

```{r tabfour, echo=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

```

## Model Evaluations

Table \@ref(tab:tabmodel) summarizes the performance metrics across the
different models fitted to the data in our investigation.

```{=tex}
\begin{table}[H]

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```
```{r tabmodel, eval=FALSE, echo=FALSE}
## Model performance
tabmodel <- model.checks %>% 
  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
tabmodel
```

## Subgroup analysis by gender

# Conclussion

## Discussion

## Limitations

Our study encountered a significant challenge with missing data,
resulting in a notable reduction in our sample size. Despite initially
having a sample of \$n=19,225\$, the final analysis was conducted on a
considerably smaller sample of \$n=1,977\$ observations. This
substantial reduction in the sample size raises concerns about the
robustness and reliability of our findings. Addressing this issue
becomes crucial to ensure the validity of our study outcomes.

In response to this challenge, we propose the implementation of multiple
imputation techniques. This approach involves generating multiple
plausible sets of missing values based on observed data patterns. By
imputing missing values using this method, we aim to enhance the
completeness of our dataset. This strategy is intended to mitigate
potential biases resulting from missing data and fortify the reliability
and accuracy of our analytical outcomes.

# References