### 2023-12-15 manuscript/paper.Rmd
final update

---
title: "Age modifies the effect of total cholesterol level on systolic blood pressure: Analysis of evidence from the 2015 - 2018 NHANES data"
thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
#subtitle: "Short title goes here!"
author:
- Amos Okutse$^1$
- Monica Colon-Vargas$^1$
- $^1$School of Public Health, Brown University, Providence, RI.
abstract: | 
  **Introduction:** Systolic blood pressure (SBP) is a precursor linked with a spectrum of critical health conditions.  This nexus between heightened SBP and these significant health adversities underscores the gravity of SBP as a predictive marker for adverse health outcomes.

  **Objective:** To explore the intricate interplay between SBP dynamics within non-obese aging populations with self-reported workday sleep duration, total cholesterol levels, and the total cholesterol-age dependence adjusting for selected potentially confounding factors.

  **Methods:** Cross-sectional data spanning two survey cycles (2015 – 2018) was obtained from the National Health and Nutrition Examination Survey (NHANES) database. We performed multivariable regressions to examine the interplay between age, total cholesterol levels, self-reported workday sleep duration, and SBP. Potentially confounding covariates adjusted for in analyses were selected using univariate regressions, best-subsets selection, and the Least Absolute Shrinkage and Selection Operator (LASSO). Model performance evaluation was based on information criteria, mean squared error (MSE), adjusted R-squared, and likelihood ratio tests.

  **Results:** Our results are based on a complete-case analysis of data on n = 1977 participants. We found a statistically significant association between workday sleep duration and SBP comparing $6 – 8$ hours sleep category to < $6$ hour sleep category [$\beta$ = -3.0; 95\% Confidence Interval (CI): -5.3,-0.67), p = 0.012] and a substantial effect modification of age on the effect of total cholesterol levels on SBP [$\beta$ = 0.11; 95\%CI: 0.07 - 0.14;  p <0.001].

  **Conclusion:** Workday sleep duration between 6-8 hours is associated with an optimal decrease in SBP. Age significantly modifies the effect of total cholesterol levels on SBP, adjusting for other potentially confounding factors.

  ***Keywords:*** *Systolic blood pressure*, *Sleep duration*, *Age*, *Total cholestrol level*, *Effect modification.*
keywords: "pandoc, r markdown, knitr"
#date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
bibliography: bibfile.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

```{r source-file, echo=FALSE, include=FALSE}
# source file 
#source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")

source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

# Introduction

The clinical importance of systolic blood pressure (SBP) stands as a
cornerstone extensively acknowledged and documented across the continuum
of medical literature's data collection [@gurven2012does;
@su2022association]. Elevated SBP stands as a precursor linked with a
spectrum of critical health conditions, encompassing cardiovascular
diseases, stroke, and renal impairment, among others. This nexus between
heightened SBP and these significant health adversities underscores the
gravity of SBP as a predictive marker for adverse health outcomes.
Noteworthy is the pivotal concern surrounding the trajectory of SBP in
the natural aging process, a trajectory observed ubiquitously across
diverse populations.

In western societies aged 40 years and above, a demonstrable pattern
emerges, revealing an approximate elevation of about 7 mmHg in SBP per decade
among individuals above 40 years. This discernible and consistent
increase in SBP with advancing age accentuates its profound impact
within the broader context of age-related health dynamics. Similarly,
while diastolic blood pressure (DBP), demonstrates a concurrent rise
correlating with age, presents a substantially lower rate of increase
when juxtaposed against the ascending trajectory observed in SBP
[@gurven2012does; @su2022association]. Thus, in light of the
significance attached to SBP dynamics within aging populations and the
intricate interplay between age, sleep, cholestrol levels and blood
pressure alterations, our investigative focus is oriented towards the
exploration of variables that exert potential influence on systolic
blood pressure (SBP). This deliberate focus aims to elucidate the
multifaceted nature of factors contributing to SBP variability, enabling
a more comprehensive understanding of its determinants within the
framework of health and aging.

```{=html}
<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized
contributor to systolic blood pressure (SBP), with individuals
exhibiting higher BMI values demonstrating a propensity for elevated
SBP. Gender also holds a discernible association with SBP dynamics.
Empirical evidence in CITE demonstrates that women typically manifest
lower SBP than men until the onset of menopause, after which women
exhibit higher SBP levels compared to men. Additionally, heightened
levels of cholesterol are closely linked to vascular damage. The
accumulation of cholesterol deposits and calcium within arteries
precipitates arterial stiffening, necessitating augmented cardiac
efforts to circulate blood, potentially culminating in elevated blood
pressure. This study endeavors to examine the interrelationships between
age and cholesterol levels, emphasizing consideration of potential
confounding variables. Table 1 comprehensively delineates the array of
variables under scrutiny, ensuring a comprehensive analysis of their
collective impact. -->
```
# Methods

## Study population

Data utilized in this paper is sourced from the National Health and
Nutrition Examination Survey (NHANES), a comprehensive nationwide survey
administered by the National Center for Health Statistics (NCHS) via the
Centers for Disease Control and Prevention (CDC). The survey assesses
the health and nutrition of the entire non-institutionalized US
population, spanning all ages and residing in all 50 states as well as
Washington D.C. As such, the survey provides a cross-sectional view of a
representative sample of the US population. Further information about
NHANES can be found at www.cdc.gov/nchs/nhanes.

## Data

Our current analyses combine the 2015 - 2018 NHANES survey cycles to
yield $n = 19225$ observations on 35 covariates. This sample size was
comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and
2017/2018 survey cycles, respectively. Analyses excluded individuals
with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055),
and body mass index (BMI) or those with BMI \> 25 kg/$m^2$ (overweight)
(n = 5521). Individuals that reported being on anti-hypertensive
medication were also excluded from further analyses (n = 2944). Our
final analyses were based on a sample of n = 1977 observations on 28
covariates.

### Outcome definition:

We defined our outcome as systolic and diastolic blood pressure. These
variables are measured by trained examiners using standardized
procedures. Given that systolic and diastolic blood pressure
measurements are taken at least four times on an individual, our
definition of these outcome is based on an average of the first three
blood pressure measurements.

### Exposures:

Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on Workdays or
workdays". We then categorized this variable into three groups, that is,
\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
group in our analysis. We also explored the association between total
cholestrol (in mmol/L) level on systolic blood pressure including
potential effect modification by other factors.

### Covariates:

```{=html}
<!-- Sleep duration on workdays was evaluated by the questionnaire with the
following questions: "Number of hours usually sleep on Workdays or
workdays". We then categorized this variable into three groups, that is,
\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
group in our analysis. -->
```
Our analyses included the following as covariates based on previous
literature [@su2022association]: Race divided into four groups as
Mexican American, white, black and other race. Alcohol consumption was
grouped into drinking, no drinking, not recorded. Smoking status as
smoking, not smoking, not recorded. Diabetes was defined as yes, no,
borderline, or not recorded. Hypertension was defined as yes, no, or not
recorded. Snorting was defined as yes, no, and not recorded. US
citizenship status was defined as citizen by birth or naturalization,
don't know, not a citizen, or refused to answer. Education level was
grouped into four categories including graduate studies, high school,
less than grade 12 or some college. Additional covariates included
marital status, gender, age, albumin, creatinine,hemoglobin, total
cholesterol (TC), aspertate aminotransferase (AST), high-density
lipoprotein (HDL), and body mass index (BMI). Details about these
variables can be found at
<https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx>. Age, albumin,
creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as
continuous variables whereas gender, alcohol consumption, diabetes,
smoking, race, hypertension, and snorting were analyzed as categorical
variables. Table \@ref{tab:tabone} highlights the variable names and
descriptions as utilized in this study.

```{=tex}
\begin{table}[H]

\caption{\label{tab:tabone}Variable descriptions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Name} & \textbf{Description}\\
\midrule
\textbf{SEQN} & \textbf{sequence number} & Respondent number\\
\textbf{SDMVPSU} & \textbf{psu} & Masked variance unit pseudo-PSU variable for variance estimation\\
\textbf{WTINT2YR} & \textbf{weights} & Full sample 2-year interview weights\\
\textbf{SDMVSTRA} & \textbf{strata} & Masked variance unit pseudo-stratum variable for variance estimation\\
\textbf{RIAGENDR} & \textbf{gender} & Respondent's number\\
\addlinespace
\textbf{RIDAGEYR} & \textbf{age (yrs)} & Respondent's age in years\\
\textbf{DMDMARTL} & \textbf{marital status} & Marital status\\
\textbf{INDFMIN2} & \textbf{income category} & Total family income (reported as a range value in dollars)\\
\textbf{RIDRETH3} & \textbf{race} & Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category\\
\textbf{DMDHHSZA} & \textbf{children <5} & Number of children aged 5 years or younger in the household\\
\addlinespace
\textbf{DMDEDUC2} & \textbf{education level} & What is the highest grade or level of school\{you have/SP has\} completed or the highest degree \{you have/s/he has\} received?\\
\textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on Workdays or workdays\\
\textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
\textbf{ALQ121} & \textbf{alcohol use} & In the past 12 months, how often did you drink any type of alcoholic beverage?\\
\addlinespace
\textbf{LBDHDDSI} & \textbf{hdl} & High density lipoprotein\\
\textbf{LBDSALSI} & \textbf{albumin} & Albumin (g/L)\\
\textbf{DIQ010} & \textbf{diabetes} & Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?\\
\textbf{BPQ020} & \textbf{hypertension} & \{Have you/Has SP\} ever been told by a doctor or other health professional that \{you/s/he\} had hypertension, also called high blood pressure?\\
\textbf{URXCRS} & \textbf{creatinine} & Creatinine, urine (umol/L)\\
\addlinespace
\textbf{SLQ040} & \textbf{snort} & In the past 12 months, how often did \{you/SP\} snort, gasp, or stop breathing while \{you were/s/he was\} asleep?\\
\textbf{LBDTCSI} & \textbf{total cholestrol levels} & Total cholesterol (mmol/L)\\
\textbf{LBXHGB} & \textbf{hemoglobin} & Hemoglobin (g/dL)\\
\textbf{LBXSASSI} & \textbf{AST} & Aspertate aminotransferase\\
\textbf{SMQ040} & \textbf{smoke} & Do you now smoke cigarettes?\\
\addlinespace
\textbf{BPXDI} & \textbf{blood pressure} & Systolic and diastolic blood pressure taken as the average of the first three measurements\\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabone, echo=FALSE, eval=FALSE}
tab.one <- vars %>% 
  kable(format = "latex", caption = "Variable descriptions", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("HOLD_position")) %>%
  row_spec(c(0), bold = TRUE) %>%
  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>%
  column_spec(3, width = "8cm")
tab.one
```

## Statistical modeling

### Variable selection based on univariate regression

Univariate regression analyses were performed on all potentially
confounding covariates excluding hypertension. Model 1 was fitted to the
data including the exposure variables as well as all potentially
confounding covariates from univariate regression analysis. Model 2 was
fitted to the same set of covariates including an interaction term
between age and total cholesterol levels. Model 3 was fitted to the data
including only covariates that had a statistically significant
association with SBP in univariate analyses excluding HDL, hemoglobin,
diabetes, survey cycle, and alcohol consumption. Model 4 included
similar covariates as Model 3 but examined, in addition, all possible
two-way interactions between the covariates. This model is essentially
overfit to the data. Model 5 included the same set of covariates as
Model 3 in addition to an interaction term between age and total
cholesterol levels.

### Variable selection based on LASSO and best subsets

Furthermore, we conducted two distinct variable selection methodologies,
similarly excluding hypertension. Model 6 was based on best subset
selection with 10-fold cross-validation to ascertain the optimal number
of subsets. Best subset selection was selected by identifying the best
model that contains a given number of predictors yielding lowest RSS.
Notably, for systolic blood pressure, the forward best subset selection
method discerned 13 variables, whereas diastolic blood pressure
exhibited 19 selected variables. However, a limitation inherent in
forward best subset selection is its inability to eliminate previously
selected features, potentially disregarding their relevance in light of
newly added variables. Foward best subset deos not outperforms shringake
methods when minimizeing prediction errors [@hastie2009elements] .
Consequently, this methodology may yield sub-optimal variable selections
due to its lack of adaptability.

In an effort to mitigate this limitation and introduce regularization,
LASSO regularization was employed. This involved a 10-fold
cross-validation process to determine the optimal penalty parameter,
$\lambda$ . For diastolic blood pressure, the $\lambda$ value associated
with the lowest mean squared error (MSE) led to the selection of 25
variables. Conversely, when considering systolic blood pressure, the
$\lambda$ value minimizing the MSE resulted in the exclusion of a sole
variable. Subsequently, a $\lambda$ value was strategically chosen to
ensure a negligible increase of no more than 1% in MSE, ultimately
leading to the inclusion of 29 variables in the model. Model 7 consisted
in variables selected by the LASSO regression. In addition, our study
incorporated an alternative model that not only incorporated the
variables identified through LASSO but also introduced an interaction
term involving age and cholesterol levels; Model 8. This extended model
aimed to explore potential combined effects between age and cholesterol
on the outcome variable.

## Model performance evaluation

All models fitted to the data in this investigation were nested and
evaluations were based on information criteria including Akaike
Information Criteria (AIC) and Bayesian Information Criteria (BIC). The
BIC is favored over the AIC due to its increasing penalty with larger
sample sizes, resulting in more stringent significance levels [@BIC].
Additionally, we also used performance measures including the mean
squared error (MSE), r-squared, and adjusted r-squared. The likelihood
ratio test (LRT) was used in competing model comparisons testing the
null hypothesis that the nested models are equivalent, and the extra
parameters in the larger model do not improve the fit significantly,
that is, these terms have effect sizes equivalent to 0.
\@ref(tab:tabmodel) summarizes the model performance metrics compared
for all models implemented in this investigation. The final model
superiority was based on likelihood ratio tests.

# Results

## Descriptive statistics

Table \@ref(tab:tabtwo) summarizes participant characteristics in the
2015 - 2018 NHANES survey cycles stratified by gender. The 2015-2016
survey cycle comprised 53.77% of the total analyzed sample whereas the
2017 - 2018 survey cycle comprised 46.23% of the total analyzed sample.
Among the participants, the proportion of males and females were 53.76%
(n = 1063) and 46.23% (n = 914), respectively. With race, the proportion
of Mexican American, White and Black were 9.16%, 34.90% and 17.10%,
respectively. Overall, the mean (SD) values for age, albumin, SBP, DBP,
hemoglobin, TC, AST, HDL, BMI were 43.79 (17.43) years, 43.04 (3.59)
g/L, 118.61 (16.41) mmHg, 69.60 (11.18) mmHg, 13.94 (1.46)g/dL, 4.81
(1.02) mmol/L, 23.47 (13.17) IU/L, 1.60 (0.43) mmol/L, and 22.08 (2.03)
kg/m2, respectively. Among the participants, 73.70% were alcohol
drinkers, 5.06% were diabetic, 7.79% were hypertensive, 13.76%
experienced snorting whereas 21.60% were smokers. Sleep duration was
divided into three groups, which were \< 6 h, 6--8 h, ≥ 8 h, each with a
proportion of 7.89%, 40.06%, 52.05%, respectively. Male and female
participants differed significantly on all characteristics except
citizenship status, and the cycle of the survey.

```{r tabtwo, echo=FALSE, message=FALSE, warning=FALSE}
tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase. SBP = Systolic blood pressure, DBP = Diastolic blood pressure.", 
                 longtable = TRUE) %>% 
  kable_styling(font_size = 8)
```

## Univariate regression analysis

Univariate regression analysis results exploring the association between
the exposures, selected potential confounding variables showed
substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 -
1.3]; p\<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p\<0.05),
albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p \< 0.05),
AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p \<0.05), and age
(years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p \<0.05) and
systolic blood pressure. Workday sleep duration between 6 and 8 hours
was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the
average systolic blood pressure compared to 6 hours of sleep. On the
other hand, more than 8 hours of sleep was associated with a decrease of
3.1 mmHg in systolic blood compared to 6 hours of sleep. Substantial
associations were noted between citizenship, education, number of
children below five years, gender, marital status, race, smoking status,
snorting, alcohol, and income.

```{r tabthree, echo=FALSE, eval=FALSE}
## univariate analysis
sbp.univariate %>% 
  as_kable_extra(format = "latex", booktabs = TRUE, digits = 2, caption = "Univariate analyses of exposures, potential confounding factors and systolic blood pressure. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase.", longtable = TRUE)
```

## Model performance evaluation

Table \@ref(tab:tabmodel) presents an overview of the performance
measures of the various models applied to our data set. Model 6,
utilizing best subset selection, revealed notably high MSE, suggesting
limited predictive efficacy. Models 1 and 2, not including variable
selection techniques, highlight the impact of incorporating the
age-cholesterol interaction term with notable reduction in MSE. Models
3, 4, and 5 included solely significant univariate regression variables.
Model 5, with the lowest AIC and second lowest BIC, demonstrates an
acceptable MSE. Model 4, with the lowest MSE and highest \$R\^2\$,
demonstrated a high predictive power, although its extensive number of
terms raise concerns of potential overfitting. Model 3, based solely on
significant univariate covariates, performed well but exhibited a higher
MSE than Models 4 and 5, suggesting the probable influence of
interaction terms. Models 7 and 8, employing LASSO, exhibited comparable
performance metrics, with Model 8 displaying reduced BIC and MSE,
attributed chiefly to the significant inclusion of an age-cholesterol
interaction term.

```{=tex}
\begin{table}[H]

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & $\text{df}^a$ & $\text{AIC}^b$ & $\text{BIC}^c$ & $\text{MSE}^d$ & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
Model 7 & 41 & 15888.38 & 16117.55 & 173.69  & 0.36 & 0.34\\
Model 8 & 42 & 15849.19 & 16083.94 & 170.11  & 0.37 & 0.36\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```
Likelihood ratio tests showed a statistically significant difference
between Model 1 and 2 with evidence of in favor of the latter
($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all
potentially confounding covariates including an
interaction between cholestrol level and age was then compared to Model
3 which adjusted for only covariates we found significant in univariate
analyses. Results here similarly showed evidence in favor of Model 2
($\chi^2_{(1)} = 7979; p>0.05$). Comparing Model 2 to Model 3 revealed a
non-statistically difference in model fit suggesting that Model 3, even
though smaller performed comparably well. Model 3 was thus selected and
compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring
its choice as the best model ($p>0.05$). Additional regression
diagnostics for Model 3 including residual versus fitted value plots,
Q-Q plots for checking the normality of residuals, Cook's distance for
outliers, residual and high leverage point plots did not reveal
substantial deviations from linear regression assumptions. Model 3 was
thus selected as the best model. Reported results are thus based on this
model.

## Multivariable Regression Analysis

Table \@ref(tab:multreg) displays the estimated coefficients derived
from a multiple regression analysis involving variables. We only show
sleep during Workdays, age, total cholesterol, and the interaction
effect between total cholesterol and age. Notably, the findings reveal
that an increase of one year in age is associated with a decrease in
systolic blood pressure (SBP) by an amount of 0.04 mmHg (95%CI [-0.22 -
0.13], p = 0.60). Conversely, an increase of one unit in total
cholesterol is linked to a significant reduction in SBP by a factor of
3.5 mmHg (95%CI [-5.2, -1.8], p = 0.01). Furthermore, the interaction
between age and total cholesterol yielded a statistically significant
coefficient estimate suggesting that as both age and total cholesterol
concurrently increase by one unit, there is an associated elevation in
SBP by a factor of 0.11 mmHg (95%CI [0.07 - 0.14], p = 0.001).
Additionally, the coefficients attributed to sleep duration bear
analogous interpretations to those observed in the univariate scenario.
Specifically, workday sleep durations ranging between 6 and 8 hours
demonstrate a significant average decrease of 3.0 mmHg (95%CI [-5.3 -
-0.67], p = 0.012) in systolic blood pressure in comparison to the
reference duration of 6 hours of sleep. Conversely, durations exceeding
8 hours exhibit a decrease of 1.4mmHg (95%CI [-3.7- 0.92] p = 0.2]) in
systolic blood pressure as compared to the 6-hour reference duration.

```{=tex}

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients based on Model 5. These results are adjusted for other significant potentially confounding factors based on univariate regression analysis.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} &  &  &  \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13) & 0.6 \\
\textbf{TC} & -3.5  & (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11  & (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

```

```{r tabmodel, eval=FALSE, echo=FALSE}
## Model performance
tabmodel <- model.checks %>% 
  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
tabmodel
```

# Discussion and conclusion

The present investigation explores the subtle interplay between age,
workday sleep patterns, cholesterol levels, and systolic blood pressure, while
accounting for various potentially confounding factors informed by literature. We conducted an exhaustive series of model
comparisons to elucidate the intricate relationship between age, total
cholesterol levels, workday sleep duration, and systolic blood pressure. In light of our findings derived from the best model based on the
performance measures in Table \@ref(tab:tabmodel), we observed that
weekday sleep durations spanning between 6 and 8 hours, as well as
durations exceeding 8 hours, exhibit a decrease in systolic blood
pressure in comparison to the reference group of individuals sleeping
for 6 hours. Furthermore, our investigation revealed a noteworthy effect
modification regarding the influence of total cholesterol levels on
systolic blood pressure when adjusted for specific confounding variables
within the studied cohort, especially in relation to age.

Our study contributes to the existing body of knowledge by
substantiating the associations uncovered in previous research
concerning sleep patterns and systolic blood pressure
[@su2022association]. This correlation underscores the significance of
sleep parameters as influential factors in understanding blood pressure
dynamics [@fuchs2020high]. In this way, the study explores factors that bear substantial weight in understanding disease risk within the aging US population. Findings from this study could inform public health policies and interventions towards healthy diet as well as sleep patterns, personalized medicine, and healthy aging promotion.

While our study used a representative sample of the US population, missing data on variables of interest in analyses presented a significant challenge. Our results are based on complete case analysis where all individuals with missing data on the variables of interest were excluded. While complete case analysis ensures that the analysis is based on comprehensive and complete information for each participant, the notable reduction in sample size has potential to result in low statistical power and bias when the assumption of missing completely at random does not hold [@baraldi2010introduction]. Future studies can address this challenge through such methods as multiple
imputation techniques. This approach involves generating multiple
plausible sets of missing values based on observed data patterns [@baraldi2010introduction].

# References